Case Study: ChIP-seq Differential Peak Calling (DiffBind)

### 2018-03-12 datasets/ChIP-seq/ChIPseq-diffbind.Rmd
Run macs2 from Rmd

For this analysis, peak calls are obtained from the ENCODE Project pipeline, and
testing is performed using _DiffBind_ and _DESeq2_.

@@

We downloaded the peak calls directly from the UCSC ENCODE portal.

```{r}
broad_url <- "http://hgdownload.soe.ucsc.edu/goldenPath/hg19/encodeDCC/wgEncodeBroadHistone/"
broad_peaks <- c("wgEncodeBroadHistoneGm12878H3k4me3StdPkRep1.narrowPeak.gz",
                 "wgEncodeBroadHistoneGm12878H3k04me3StdPkRep2V2.narrowPeak.gz",
                 "wgEncodeBroadHistoneK562H3k4me3StdPkRep1.narrowPeak.gz",
                 "wgEncodeBroadHistoneK562H3k4me3StdPkRep2.narrowPeak.gz")

for (i_peak in broad_peaks) {
    if (!file.exists(file.path(datdir, i_peak))) {
        download.file(paste0(broad_url, i_peak),
                      destfile = file.path(datdir, i_peak))
    }
}

uw_url <- "http://hgdownload.soe.ucsc.edu/goldenPath/hg19/encodeDCC/wgEncodeUwHistone/"
uw_peaks <- c("wgEncodeUwHistoneGm12878H3k4me3StdPkRep1.narrowPeak.gz",
              "wgEncodeUwHistoneGm12878H3k4me3StdPkRep2.narrowPeak.gz",
              "wgEncodeUwHistoneK562H3k4me3StdPkRep1.narrowPeak.gz",
              "wgEncodeUwHistoneK562H3k4me3StdPkRep2.narrowPeak.gz")

for (i_peak in uw_peaks) {
    if (!file.exists(file.path(datdir, i_peak))) {
        download.file(paste0(uw_url, i_peak),
                      destfile = file.path(datdir, i_peak))
    }
}
```

@@

```{r}
peakfiles <- file.path(datdir, c(broad_peaks, uw_peaks))
labs <- gsub("wgEncode(.*)Histone.*", "\\1", basename(peakfiles))
cells <- gsub("wgEncode.*Histone(.*)H3k.*", "\\1", basename(peakfiles))

meta <- data.frame(cellline = cells, lab = labs, file = peakfiles,
                   stringsAsFactors = FALSE)
```

>>>>

For this analysis, peak calls are made using MACS2 on reads aligned using the _Rsubread_ package,
and testing is performed using _DiffBind_ and _DESeq2_.

@@

@@

```{r}
## create metadata table
sortedfiles <- list.files(datdir, pattern = "\\.sorted\\.bam$", full.names = TRUE)
labs <- gsub("wgEncode(.*)Histone.*", "\\1", basename(sortedfiles))
cells <- gsub("wgEncode.*Histone(.*)H3k.*", "\\1", basename(sortedfiles))
meta <- data.frame(cellline = cells, lab = labs, sortedfile = sortedfiles,
                   peakfile = gsub("\\.bam$", "_peaks.narrowPeak", sortedfiles),
                   stringsAsFactors = FALSE)
```

We call peaks using MACS using reads aligned in `ChIPseq-csaw.Rmd`.

```{r}
for (i in seq_len(nrow(meta))) {
    i_bam <- meta$sortedfile[i]
    i_peak <- meta$peakfile[i]
    if (!file.exists(i_peak)) {
        code <- sprintf("macs2 callpeak -t %s -f BAM -g hs -n %s -q 0.1",
                        i_bam, gsub("\\.bam", "", i_bam))
        code <- system(code)
        stopifnot(code == 0L)
    }
}
```
